---
title: "Hackday 9th Oct"
author: "ruby"
date: "2024-10-09"
output: pdf_document
---
#set up
```{r}
library(data.table)
library(ggplot2)
library(patchwork)
library(lubridate)
setwd("~/Documents/Documents - Ruby's MacBook Air/RAC Work/MotoringIndex/RubyMotoringIndex")
```

#Read files
```{r}
DftStats1 <- fread("dft-road-casualty-statistics-casualty-1979-latest-published-year.csv")
AdjustedStats <- fread("dft-road-casualty-statistics-casualty-adjustment-lookup_2004-latest-published-year.csv")
TotalMilesDriven <- fread("~/Documents/Documents - Ruby's MacBook Air/RAC Work/MotoringIndex/tra0101-miles-by-vehicle-type.csv")
```
```{r}
DftStats2 <- merge(dfst
                   )
```


```{r}
DftStats1 <- DftStats1[, c("sex_of_casualty", "age_of_casualty", "age_band_of_casualty", "pedestrian_location", "pedestrian_movement", "pedestrian_road_maintenance_worker", "casualty_home_area_type", "casualty_distance_banding", "casualty_imd_decile", "casualty_class", "car_passenger", "bus_or_coach_passenger", "casualty_type", "lsoa_of_casualty", "accident_reference", "vehicle_reference") := NULL]
AdjustedStats <- AdjustedStats[, c("accident_reference", "vehicle_reference", "casualty_reference"):= NULL]
```


#Formatting
```{r}
DftStats2004 <- DftStats1[accident_year >= 2004,]
DftStats2004 <- DftStats2004[casualty_severity < 3,]
DftStats1[, accident_reference, accident_year]
```

```{r}
AdjustedStats <- merge(AdjustedStats, DftStats1, by = c("accident_index", ), all = TRUE)
rm(DftStats2004)
```

```{r}
CasualtybyYear <- DftStats1[,.N, keyby = c("accident_year", "casualty_severity")]
xyz <- subset(CasualtybyYear, CasualtybyYear$casualty_severity == "1")
abc <- subset(CasualtybyYear, CasualtybyYear$casualty_severity == "2")
CasualtybyYear <- merge(xyz, abc, by = "accident_year", all = TRUE)
setnames(CasualtybyYear, c("N.x", "N.y"), c("Fatalities", "Serious Injuries"))
CasualtybyYear[, c("casualty_severity.x", "casualty_severity.y") := NULL]
rm(xyz)
rm(abc)
CasualtybyYear <- CasualtybyYear[, "KSI" := Fatalities + `Serious Injuries`]
```

```{r}
TotalMilesDriven <- TotalMilesDriven[, Months := month(`Year ending [note 1]`, label = TRUE)]
TotalMilesDriven <- subset(TotalMilesDriven, TotalMilesDriven$Months == "Dec")
TotalMilesDriven <- TotalMilesDriven[,`Year ending [note 1]` := year(`Year ending [note 1]`)]
TotalMilesDriven[, c("Cars and Taxis", "Light Commercial Vehicles [note 3]", "Heavy Goods Vehicles [note 4]", "Other Vehicles [note 5]", "Months") := NULL]
TotalMilesDriven <- merge(TotalMilesDriven, CasualtybyYear, by.x = "Year ending [note 1]" , by.y = "accident_year", all= TRUE)
TotalMilesDriven <- TotalMilesDriven[16:45]
TotalMilesDriven <- TotalMilesDriven[, FatalityRate := Fatalities / `All Motor Vehicles`]
TotalMilesDriven <- TotalMilesDriven[, CasualtyRate := `Serious Injuries` / `All Motor Vehicles`]
TotalMilesDriven <- TotalMilesDriven[, KSIRate := KSI / `All Motor Vehicles`]
```

#Graphs
```{r}
ggplot() +
  geom_line(data = CasualtybyYear, aes(x = accident_year, y= Fatalities ,colour = "Fatality (killed)") ) +
  geom_line(data = CasualtybyYear, aes(x = accident_year, y = `Serious Injuries` , colour = "Casualty (Serious injury)")) +
  theme_classic() + ylab("Totals") + scale_color_manual("Casuality severity levels", values = CasualtyColours) + xlab("End of year")
```


```{r}
ggplot(data= CasualtybyYear, aes(x= accident_year, y = KSI, colour = "Fatality or casualty (killed or seriously injured)")) + geom_line()+ theme_classic() + ggtitle("Annual KSI's") + ylab("Totals") + 
  scale_color_manual("Casuality severity levels", values = CasualtyColours) + xlab("End of year")
```
#Graphs by miles driven
```{r}
ggplot() + 
  geom_line(data = TotalMilesDriven, aes(y = KSIRate, x= `Year ending [note 1]`, colour = "Fatality or casualty (killed or seriously injured)"))+
  theme_classic() + 
  ggtitle("Fatalities and serious injuries per one billion motor vehicle miles") + ylab("Rate") + 
  scale_color_manual("Casuality severity levels", values = CasualtyColours) +
  theme(legend.position = "bottom", legend.title.position = "top", legend.text.position = "bottom") + xlab("End of year") 
```
KSIPlot <- 
```{r}
ggplot() + 
  geom_line(data = TotalMilesDriven, aes(y = FatalityRate, x= `Year ending [note 1]`, colour = "Fatality (killed)", linetype = "Fatality (killed)")) + 
  geom_line(data = TotalMilesDriven, aes(y = CasualtyRate, x= `Year ending [note 1]`, colour = "Casualty (Serious injury)", linetype = "Casualty (Serious injury)")) + 
  geom_line(data = TotalMilesDriven, aes(y = KSIRate, x= `Year ending [note 1]`, colour = "Fatality or casualty (killed or seriously injured)", linetype = "Fatality or casualty (killed or seriously injured)"))+
  theme_classic() + 
  ggtitle("Fatalities and serious injuries per one billion motor vehicle miles travelled") + ylab("Rate") + 
  scale_color_manual("Casuality Severity Levels", values = CasualtyColours) + xlab("End of year") + scale_linetype_manual("Casuality Severity Levels", values = CasualtyLines)+
  theme(legend.position = "bottom", legend.title.position = "top", legend.text.position = "bottom")

```
ggsave("KSIPlot.pdf")
#Notes
PlotA <- (plot1 + plot2)
PlotB <- (plot3 + plot4)

ggsave("PlotA.pdf")
ggsave("PlotB.pdf")
ggsave("plot5.pdf")

CombinedB <- ((Plot2022 + Plot2023) + plot_annotation(title = "Pass rates for fuel types and age of vehicle"))
MonthlyRPI2003 <- MonthlyRPI2003[, Title := seq(as.Date("2003-06-01"), as.Date("2024-08-01"), by = "mon")]
WeeklyFuelData <- WeeklyFuelData[, MonthYear := paste0(month(Date, label = TRUE), year(Date))]

1 = fatality, 2 = serious